Cominuscule subvarieties of flag varieties

### 2025-10-24 ms.tex
Add files via upload

\usepackage{lmodern}
\usepackage{cfr-lm}
\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenx}
\usepackage{lineno}
\linenumbers
\usepackage{siunitx}
\usepackage{xparse}
\usepackage{standalone}
\usepackage{mathtext}
\usepackage[english]{babel}
\usepackage{hyperref}%
\hypersetup{
  colorlinks   = true,  %Colours links instead of ugly boxes
  urlcolor     = black, %Colour for external hyperlinks
  linkcolor    = black, %Colour of internal links
  citecolor    = black  %Colour of citations
}
\usepackage[kerning=true,tracking=true]{microtype}
\usepackage{xspace}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{mathabx}
\usepackage{braket}
\usepackage{cool}
\usepackage{varioref}
\usepackage{xstring}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{longtable}
\usepackage[longtable]{multirow}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{varwidth}
\usepackage{tensor}
\usepackage{scalerel}
\usepackage{lie-hasse}
\tikzset{
	background rectangle/.style={
		shade,
		top color=olive!20,
		bottom color=white,
		draw=olive!15,
		very thick,
		rounded corners}}
\usepackage{rank-2-roots}
\usepackage{tikz-cd}
\usepackage{colortbl}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\theoremstyle{remark}
\newtheorem{conjecture}{Conjecture}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
%\newtheorem{remark}{Remark}
%\newtheorem{Comment}{Comment}
%\newcounter{remarkCounter}
%\setcounter{remarkCounter}{1}
%\NewDocumentCommand\todo{m}
%{%
%    \par\noindent{\textcolor{blue}{\textbf{Comment \#\arabic{remarkCounter}. } #1}}
%    \par
%    \stepcounter{remarkCounter}
%}%
%\NewDocumentCommand\finish{}{\todo{Finish this.}}

@@

\NewDocumentEnvironment{longtabl}{m}%
\renewcommand{\arraystretch}{2}%
\begin{longtable}{#1}%
\end{longtable}%
\renewcommand{\arraystretch}{1}%

@@

While we cannot draw a flag variety, or even its associated root system (except in low dimensions), we can always draw its Hasse diagram.

@@

Since cominiscule varieties are simpler than other flag varieties in many ways, we hope that these cominiscule varieties will shine light on their ambient flag varieties.
We give some evidence for their significance by demonstrating that the associated cominiscule subvariety of any irreducible flag variety is the unique submanifold of maximal symmetry group among all submanifolds satisfying a certain open condition on derivatives.

@@

Each Cartan subgroup of the projective transformations of the plane consists of those which preserve three points in general position:

@@

Hence the associated cominiscule is invariant under a Cartan subgroup, and conversely there are finitely many associated cominiscules invariant under any given Cartan subgroup.
The projective transformations preserving the point \(p_0\) and line \(\ell_0\) act transitively on the associated cominiscule, moving the points \(p\) of the line \(\ell_0\).

@@

for which 

@@

A \emph{flag variety} \((X,G)\), also called a \emph{generalized flag variety} or a \emph{rational homogeneous variety}, is a complex projective variety \(X\) acted on transitively and holomorphically by a connected complex semisimple Lie group \(G\) \cite{Humphreys:1975} pp. 134--135.

@@

Each flag variety is determined uniquely, up to finite central extension of \(G\) and up to isomorphism, by the Dynkin diagram of \(G\) decorated with \(\dynkin{A}{*}\) on each compact simple root and \(\dynkin{A}{x}\) on each noncompact simple root \cite{Borel:1991} p. 197 Proposition 14.18, \cite{Humphreys:1975} p. 197 Theorem${}'$.

@@

The center of any complex semisimple Lie group \(G\) lies in every maximal torus, so in every Cartan subgroup \cite{Borel:1991} p. 220, so in every parabolic subgroup, so acts trivially on every flag variety.

@@

This occurs just when there is a compact subgroup \(K\subseteq G\) so that \(\pr{X,K}\) is a compact Hermitian symmetric space \cite{Kostant:1961} p. 379 Proposition 8.2, \cite{Baston/Eastwood:1989} p. 26.

@@

Every effective cominuscule variety is a product of the following irreducible effective cominuscule varieties \cite{Kobayashi/Nagano:1964} theorem 1 p. 401:\par\noindent{}%

@@

component of pace of null $r$-planes in $\C[2r]$ \\

@@

Every complex linear algebraic group \(G\) has a \emph{unipotent radical}, the unique maximal unipotent normal subgroup, which is a closed complex linear algebraic subgroup \cite{Borel:1991} p. 85 Theorem 4.5, p. 86 Theorem 4.7, p. 157 11.21.

A complex linear algebraic group is \emph{reductive} if it contains a Zariski dense compact subgroup \cite{Vinberg:1994} p. 142 Theorem 2.7, Theorem 2.8.
Every complex linear algebraic group \(G\) has a \emph{reductive Levi factor}, i.e. a maximally reductive complex linear algebraic subgroup,  unique up to conjugacy, so that \(G\) is a semidirect product of its reductive Levi factor and its unipotent radical \cite{Borel:1991} p. 158 11.22.
Any Cartan subgroup is therefore a subgroup of the Levi factor, after perhaps a conjugacy.
Every compact subgroup lies in a maximal compact subgroup, which is unique up to conjugacy, so lies in the Levi factor up to conjugacy \cite{Hilgert.Neeb:2012} p. 531 Theorem 14.1.3.
After perhaps extending by some finite group of order a power of \(2\), the Weyl group embeds in \(G\) as a finite subgroup \cite{Tits:1966}, hence compact, so this group lies in the Levi factor up to conjugacy.

@@

A Zariski closed subgroup \(P\subseteq G\) of a connected linear algebraic group \(G\) is \emph{parabolic} if \(X:=G/P\) is a projective variety, and this occurs just when \((X,L)\) is a flag variety for a semisimple Levi factor \(L\subseteq G\), and this occurs just when \(P\) contains a Borel subgroup (i.e. a maximal connected solvable subgroup) \cite{Borel:1991} p. 148.
Every parabolic subgroup is connected p. 197 Proposition 14.18.
The unipotent radical of \(P\) is denoted \(\GN\subseteq P\), and a maximal reductive Levi factor is denoted \(\GZ\subseteq P\), so \(P=\GZ\ltimes\GN\).
(This is potentially confusing; the reader might expect to write these as \(P_+\) and \(P_0\) since they lie in \(P\), but this notation is standard \cite{Cap/Slovak:2009} p. 293 theorem 3.2.1, and due to the presence of the grading of the Lie algebra of \(G\) which we will define.)
A flag variety is cominuscule just when \(\GN\) is abelian \cite{Cap/Slovak:2009} p. 296 \S{}3.2.3.

@@

Two parabolic subgroups \(P,\op{P}\subseteq G\) of a complex semisimple Lie group are \emph{opposite} if \(P\cap\op{P}\) is a maximal reductive Levi subgroup of both \(P\) and \(\op{P}\).
All Borel subgroups of \(G\) are conjugate \cite{Borel:1991} p. 147 chapter IV 11.1, each containing a Cartan subgroup, hence the Lie algebra of \(P\) is a sum of root spaces with the Cartan subalgebra.
Every automorphism of a root system arises from an automorphism of the associated semisimple Lie group.

@@

An open subset of \(G\) consists of elements uniquely expressed as a product \(p,q\in P\times G_-\mapsto pq\in G\) \cite{Cap/Slovak:2009} p. 294, \cite{Borel:1991} p. 198 Proposition 14.21.
Every root system also has an automorphism, traditionally called \(w_0\), which belongs to the Weyl group and which interchanges the positive and negative roots of a root system \cite{Cap/Slovak:2009} p. 323-324; it is the unique element of the Weyl group of minimum length.
Note that \(w_0\) might not reverse the signs of simple roots \cite{Cap/Slovak:2009} p. 324.

@@

\subsection{Definition of the associated cominiscule}

@@

\subsection{Finding the associated cominiscule subvariety}

@@

\subsection{Why the associated cominiscule subvariety matters}
We will see \vpageref{subsection:Freedom} that the associated cominiscule subvariety of an irreducible flag variety \((X,G)\) satisfies an open condition on its tangent spaces, which we call \emph{freedom}.
We will see that the symmetry group of the associated cominiscule subvariety has strictly maximal dimension among symmetry groups of subvarieties of \(X\) with free smooth locus.
We will see that all free smooth subvarieties are homogeneous, evidence for the conjecture that the associated cominiscule subvariety is the unique free smooth subvariety.

@@

With \(\dynkin{A}{o}\) denoting a node which could be either a \(\dynkin{A}{x}\) or \(\dynkin{A}{*}\), the associated cominuscule subvarieties are (allowing some redundacy where it might clarify):

@@

%\dynkinBrace[r-1]{1}{5}

@@

\Flag{\dynkin C{*x}}{\dynkin C{*x}}
\Flag{\dynkin C{**x}}{\dynkin C{**x}}
\Flag{\dynkin[parabolic=16]{C}{}}{\dynkin[parabolic=16]{C}{}}%
\Flag{\dynkin C{xo}}{\dynkin A{x}}
\Flag{\dynkin C{xoo}}{\dynkin A{x}}
\Flag{\dynkin C{xo.ooo}}{\dynkin{A}{x}}%
\Flag{\dynkin C{*xo}}{\dynkin C{*x}}
\Flag{\dynkin C{**xo}}{\dynkin C{**x}}
\Flag{\dynkin C{*.*xo}}{\begin{dynkinDiagram}{C}{**.*x}\dynkinBrace[r-1]{1}{4}\end{dynkinDiagram}}
\LastFlag{7}{C_r}{%

@@

\dynkinBrace[\ell\ge 2]{1}{3}

@@

@@

\Flag{\dynkin{D}{oxooooo}}{\dynkin{A}{x}}

@@

For a flag variety \(X=G/P\), the root system is also \emph{\(P\)-graded} by \(\sum n_i\), but summing only over the noncompact (crossed) simple roots \cite{Cap/Slovak:2009} p. 292.
The \(P\)-grade is also called the \emph{\(P\)-height} \cite{Cap/Slovak:2009} p. 292.
A root is \emph{\(P\)-maximal} if it has maximal \(P\)-grade in its irreducible factor.
The \emph{box} is the set of \(P\)-maximal roots, terminology which roughly follows what \cite{Buch.Chaput.Mihalcea.Perrin:2018}, \cite{Lam.Williams:2008} p. 57 might perhaps call the \emph{maximal box}, by analogy with Young tableaux.
It is easy to see (see the proof~\vpageref{page:associated.cominuscule} of lemma~\ref{lemma:associated.cominuscule}) that the box generates the root system of \(\breve{G}\).

@@

The unipotent radical \(G_+\subseteq P\) has Lie algebra \(\LieG_+\subseteq\LieP\) the sum of the root vectors of the positively graded roots \cite{Borel:1991} p. 197 Proposition 14.18, \cite{Knapp:2002} p. 482.

@@

The sum of the Cartan subalgebra with the sum of the root soaces of the zero graded roots is the semisimple factor of the Levi factor \(\LieGZ\) of \(\LieP\) \cite{Borel:1991} p. 197 Proposition 14.18.

@@

We can and will pick the hyperplane to be the zero locus of a real linear function taking on value zero on uncrossed simple roots and one on crossed simple roots \cite{Cap/Slovak:2009} p. 239 proposition 3.1.2.

@@

Under the heading  \(\breve\LieG\), we draw the roots of the symmetry Lie algebra of the associated cominiscule, and under the heading \(\LieG'\) the roots of the automorphism Lie algebra.

@@

Similarly, the associated cominiscule of any adjoint variety is \(\dynkin[parabolic=1]{A}{1}=(\Proj{1},\PSL{2})\), since the adjoint varieties are precisely those with invariant contact structures, which arise from a hyperplane in each tangent space, invariant under the parabolic, hence a sum of root spaces, so a single root, negative for the parabolic, doesn't have its root vector in this hyperplane, i.e. a single root is at higher weight.
So the associated cominiscule is complementary to the hyperplane, one dimensional, hence a projective line.

@@

Recall that every root system has a unique highest root, highest in the sense of the positive roots, i.e. the highest weight of the adjoint representation \cite{Serre:2001} p. 61 Theorem 3.
Every root can be brought to the highest root by successively adding simple roots, passing through a sequence of roots \cite{Fulton/Harris:1991} pp. 330--331 \S{21.3}, \cite{Humphreys:1978} p. 56, \cite{Knutson:2022}, \cite{Serre:2001} p. 46 2.12, \cite{Serre:2001} p. 58 Theorem 3.
As this process can only raise the \(P\)-height, the highest root is \(P\)-maximal.
Every \(P\)-maximal root can thus be brought to the highest root by adding only \(P\)-compact simple roots.
So the box becomes a connected graph, with the highest root as one vertex.
The vector space \(\LieZ'\) is therefore an irreducible \(\GZ\)-module \cite{Knapp:2002} p. 332 proposition 5.105.

Pick a root vector \(e_{\aa}\in\LieZ\) for a root \(\aa\in B\).
We need to prove that \(\aa\in B'\) i.e. \(\aa\) is a box root, i.e. \(P\)-maximal.
Suppose that \(\aa+\bb\) is a root, for some root \(\bb\) of positive \(P\)-height.
So then \(0\ne\lb{e_{\aa}}{e_{\bb}}\in\gr[\aa+\bb]\LieG\) for some root vector \(e_{\bb}\) of \(\bb\) \cite{Serre:2001} p. 46 2.12.
But then \(e_{\aa}\) is not in \(\LieZ\), i.e. does not centralize \(\LieG_+\).
But the highest root is \(\aa\) or is of the form \(\aa+\bb\) for some positive root \(\bb\).
Hence \(\bb\) is \(P\)-compact, i.e. \(\aa\) is \(P\)-maximal.
So \(\aa\) is a box root: \(\aa\in B'\).

@@

By Langlands decomposition \cite{Knapp:2002} p. 482,  \(G,P,X,G_+,G_0\) are connected.

@@

Explicitly calculating out in a Chevalley basis that each element
g=e^{\sum t_{\alpha} e_{\alpha}}\in G_+
acts on an element \(e_{\beta}\in G_+\) by
\Ad_g e_{\beta}
expanding out into a sum with entirely positive coefficients unless \(\log g\) is a sum of \(P\)-maximal roots, so the center \(Z=Z_{G_+}\) of \(G_+\) consists precisely of the exponentials of elements of
\LieZ=\bigoplus\LieG_{\alpha},
the sum being over the box, i.e. over the \(P\)-maximal roots, so \(Z\) is connected.
By definition of \(\breve{G}\), \(\breve{G}:=\left<Z,aZ\right>\) is connected, since \(Z\) is connected. 

@@

%Two associated cominiscule subvarieties are tangent at a point just when they are equal.

@@

The vector spaces \(\LieZ\) and \(a\LieZ\) are irreducible \(\GZ\)-modules \cite{Knapp:2002} p. 332 proposition 5.105.
Hence \(\breve\LieG\) is a \(\GZ\)-module.
As \(\GZ\) is reductive, \(\breve\LieG\) is a direct sum of irreducible \(\GZ\)-modules:
\[
\breve\LieG=\LieZ\oplus{\breve\LieG}_0\oplus a\LieZ.
\]

Suppose that \(\aa\) is a \(P\)-compact root, i.e. a root of \(\GZ\).
Reflection in \(\aa\) is carried out by an element of the Weyl group of \(\GZ\), and so preserves the \(P\)-grading.
So if \(\bb\) is any \(P\)-maximal root, then so is its \(\aa\)-reflection.
In other words, reflection in \(\aa\) preserves the box.
Reflection in \(\aa\) moves \(\bb\) along an \(\aa\)-root string.

@@

By definition \(P'\) preserves \(\breve\LieG+\LieP\).
Pick a root \(\mx\bb\) in the box, not perpendicular to our \(P\)-submaximal \(P\)-positive root \(\aa\).
Note that if two roots have more than a right angle between them, then their sum is a root and their root vectors have nonzero Lie bracket \cite{Serre:2001} p. 29.
Suppose that \(\aa\) is \(P\)-positive.
The root system generated by \(\aa\) and \(\mn\bb:=-\mx\bb\) is of rank \(2\), and our \(P\)-grading grades that rank \(2\) root system, so we can examine in our pictures above every possible case.
We see that the root vectors of \(\aa\) and \(\mn\bb\) have more than a right angle between them and that their sum is a root.
Since their \(P\)-grading is at least that of \(\mn\bb\), \(P\)-negative, the Lie bracket of their root vectors is not in \(\LieP\), but also is not \(P\)-minimal, so not in \(\breve\LieG\).
Hence their root vectors bracket out of \(\breve\LieG+\LieP\), 
Hence the Lie algebra \(\LieG'\) of \(G'\) contains no such.

@@

The \emph{Hasse diagram} draws dots on the plane, one for each positive root and a line from each root to each of its successors, labelled by the number of the simple root by which they differ.

@@

For example, the Hasse diagram of \(C_4\) is

@@

Compare \dynkin F4 to \dynkin F{***x}:

>>>>

\usepackage{lmodern}
\usepackage{cfr-lm}
\usepackage[T2A,T1]{fontenc}
\usepackage[utf8]{inputenx}
%\usepackage{lineno}
%\linenumbers
\usepackage{siunitx}
\usepackage{xparse}
\usepackage{standalone}
\usepackage{mathtext}
\usepackage[english]{babel}
\usepackage{hyperref}%
\hypersetup{
  colorlinks   = true,  %Colours links instead of ugly boxes
  urlcolor     = black, %Colour for external hyperlinks
  linkcolor    = black, %Colour of internal links
  citecolor    = black  %Colour of citations
}
\usepackage[kerning=true,tracking=true]{microtype}
\usepackage{xspace}
\usepackage{amsfonts}
\usepackage{verbatim}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{mathabx}
\usepackage{braket}
\usepackage{cool}
\usepackage{varioref}
\usepackage{xstring}
\usepackage{array}
\usepackage{ragged2e}
\usepackage{longtable}
\usepackage[longtable]{multirow}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{varwidth}
\usepackage{tensor}
\usepackage{scalerel}
\usepackage{lie-hasse}
\tikzset{
	background rectangle/.style={
		shade,
		top color=olive!20,
		bottom color=white,
		draw=olive!15,
		very thick,
		rounded corners}}
\usepackage{rank-2-roots}
\usepackage{tikz-cd}
\usepackage{colortbl}
\newtheorem{theorem}{Theorem}
\newtheorem{corollary}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\theoremstyle{remark}
\newtheorem{conjecture}{Conjecture}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{remark}{Remark}
\newtheorem{Comment}{Comment}
\newcounter{remarkCounter}
\setcounter{remarkCounter}{1}
\NewDocumentCommand\todo{m}
{%
    \par\noindent{\textcolor{blue}{\textbf{Comment \#\arabic{remarkCounter}. } #1}}
    \par
    \stepcounter{remarkCounter}
}%
\NewDocumentCommand\finish{}{\todo{Finish this.}}

@@

\NewDocumentEnvironment{longtabl}{m}%
{%
\renewcommand{\arraystretch}{2}%
\begin{longtable}{#1}%
}%
{%
\end{longtable}%
\renewcommand{\arraystretch}{1}%
}%
\tikzset{
  snake left/.style={
    rounded corners,
    to path={
      let \p1 = (\tikztostart.east),
          \p2 = (\tikztotarget.west),
          \p3 = ($(\p1)!0.5!(\p2)$),
          \n1 = {8pt} 
      in
      (\p1)
      -- (\x1 + \n1, \y1)
      -- (\x1 + \n1, \y3)
      -- (\x2 - \n1, \y3) \tikztonodes
      -- (\x2 - \n1, \y2)
      -- (\p2)
    }
  }
}

@@

We cannot draw a flag variety, or even its associated root system (except in low dimensions), but we can always draw its Hasse diagram.

@@

Cominuscule varieties are simpler than other flag varieties in many ways; we hope that these cominuscule varieties will shine light on their ambient flag varieties.
We give some evidence for their significance by demonstrating that the associated cominuscule subvariety of any irreducible flag variety is the unique submanifold of maximal symmetry group among all submanifolds satisfying a certain open condition on derivatives.

@@

Each Cartan subgroup of the projective transformations of the plane consists of those projective transformations which preserve three points in general position:

@@

Hence the associated cominuscule is invariant under a Cartan subgroup, and conversely there are three associated cominuscules invariant under any given Cartan subgroup.
The projective transformations preserving the point \(p_0\) and line \(\ell_0\) act transitively on the associated cominuscule, moving the points \(p\) of the line \(\ell_0\).

@@

obtained by taking any linear subspace \(W_1\subseteq V_1\oplus V_k\) of the same dimension as \(V_1\), and then taking
W_2:=W_1\oplus V_2, W_3:=W_1\oplus V_2\oplus V_3,\dots,
W_k:=W_1\oplus V_2\oplus\dots\oplus V_k.
In other words,

@@

A \emph{flag variety} \((X,G)\), also called a \emph{generalized flag variety} or a \emph{rational homogeneous variety}, is a complex projective variety \(X\) acted on transitively and holomorphically by a connected complex semisimple Lie group \(G\) \cite{Humphreys:1975} pp.~134--135.

@@

Each flag variety is determined uniquely, up to finite central extension of \(G\) and up to isomorphism, by the Dynkin diagram of \(G\) decorated with \(\dynkin{A}{*}\) on each compact simple root and \(\dynkin{A}{x}\) on each noncompact simple root \cite{Borel:1991} p.~197 Proposition 14.18, \cite{Humphreys:1975} p.~197 Theorem${}'$.

@@

The center of any complex semisimple Lie group \(G\) lies in every maximal torus, so in every Cartan subgroup \cite{Borel:1991} p.~220, so in every parabolic subgroup, so acts trivially on every flag variety.

@@

This occurs just when there is a compact subgroup \(K\subseteq G\) so that \(\pr{X,K}\) is a compact Hermitian symmetric space \cite{Kostant:1961} p.~379 Proposition 8.2, \cite{Baston/Eastwood:1989} p.~26.

@@

Every effective cominuscule variety is a product of the following irreducible effective cominuscule varieties \cite{Baston/Eastwood:1989} p.~26, Example~3.1.10, \cite{Kobayashi/Nagano:1964} theorem~1 p.~401:
%\par\noindent{}%

@@

@@

Every complex linear algebraic group \(G\) has a \emph{unipotent radical}, the unique maximal unipotent normal subgroup, which is a closed complex linear algebraic subgroup \cite{Borel:1991} p.~85 Theorem~4.5, p.~86 Theorem~4.7, p.~157 11.21.

A complex linear algebraic group is \emph{reductive} if it contains a Zariski dense compact subgroup \cite{Vinberg:1994} p.~142 Theorem~2.7, Theorem~2.8.
Every complex linear algebraic group \(G\) has a \emph{reductive Levi factor}, i.e. a maximally reductive complex linear algebraic subgroup,  unique up to conjugacy, so that \(G\) is a semidirect product of its reductive Levi factor and its unipotent radical \cite{Borel:1991} p.~158 11.22.
Any Cartan subgroup is therefore a subgroup of the reductive Levi factor, after perhaps a conjugacy.
Every compact subgroup lies in a maximal compact subgroup, which is unique up to conjugacy, so lies in the reductive Levi factor up to conjugacy \cite{Hilgert.Neeb:2012} p.~531 Theorem~14.1.3.
After perhaps extending by some finite group of order a power of \(2\), the Weyl group embeds in \(G\) as a finite subgroup \cite{Tits:1966}, hence compact, so this group lies in the reductive Levi factor up to conjugacy.

@@

A Zariski closed subgroup \(P\subseteq G\) of a connected linear algebraic group \(G\) is \emph{parabolic} if \(X:=G/P\) is a projective variety, and this occurs just when \((X,L)\) is a flag variety for a semisimple Levi factor \(L\subseteq G\), and this occurs just when \(P\) contains a Borel subgroup (i.e. a maximal connected solvable subgroup) \cite{Borel:1991} p.~148.
Every parabolic subgroup is connected \cite{Borel:1991} p.~197 Proposition 14.18.
The unipotent radical of \(P\) is denoted \(\GN\subseteq P\), and a reductive Levi factor is denoted \(\GZ\subseteq P\), so \(P=\GZ\ltimes\GN\).
(This is potentially confusing; the reader might expect to write these as \(P_+\) and \(P_0\) since they lie in \(P\), but this notation is standard \cite{Cap/Slovak:2009} p.~293 theorem 3.2.1, and due to the presence of the grading of the Lie algebra of \(G\) which we will define.)
A flag variety is cominuscule just when \(\GN\) is abelian \cite{Cap/Slovak:2009} p.~296 \S{}3.2.3.

@@

Two parabolic subgroups \(P,\op{P}\subseteq G\) of a complex semisimple Lie group are \emph{opposite} if \(P\cap\op{P}\) is a reductive Levi factor of both \(P\) and \(\op{P}\).
All Borel subgroups of \(G\) are conjugate \cite{Borel:1991} p.~147 chapter IV 11.1, each containing a Cartan subgroup, hence the Lie algebra of \(P\) is the sum of (1) the Cartan subalgebra with (2) various root spaces.
Every automorphism of a root system arises from an automorphism of the associated semisimple Lie group \cite{Fulton/Harris:1991} p.~498 Proposition D.40, \cite{Humphreys:1978}, p.~87, \S16.5.

@@

An open subset of \(G\) consists of elements uniquely expressed as a product \(p,q\in P\times G_-\mapsto pq\in G\) \cite{Cap/Slovak:2009} p.~294, \cite{Borel:1991} p.~198 Proposition 14.21.
Every root system also has an automorphism, traditionally called \(w_0\), which belongs to the Weyl group and which interchanges the positive and negative roots of a root system \cite{Cap/Slovak:2009} pp.~323-324; it is the unique element of the Weyl group of maximum length.
Note that \(w_0\) might not reverse the signs of simple roots \cite{Cap/Slovak:2009} p.~324.

@@

\subsection{Definition of the associated cominuscule}

@@

\subsection{Finding the associated cominuscule subvariety}

@@

\subsection{Why the associated cominuscule subvariety matters}
We will see \vpageref{subsection:Freedom} that the associated cominuscule subvariety of an irreducible flag variety \((X,G)\) satisfies an open condition on its tangent spaces, which we call \emph{freedom}.
We will see that the symmetry group of the associated cominuscule subvariety has strictly maximal dimension among symmetry groups of subvarieties of \(X\) with free smooth locus.
We will see that all free smooth subvarieties are homogeneous, evidence for the conjecture that the associated cominuscule subvariety is the unique free smooth subvariety.

@@

With \(\dynkin{A}{o}\) denoting a node which could be either a \(\dynkin{A}{x}\) or \(\dynkin{A}{*}\), each irreducible flag variety has associated cominuscule subvariety (allowing some redundacy where it might clarify):

@@

@@

%\Flag{\dynkin C{*x}}{\dynkin C{*x}}
%\Flag{\dynkin C{**x}}{\dynkin C{**x}}
%\Flag{\dynkin[parabolic=16]{C}{}}{\dynkin[parabolic=16]{C}{}}%
%\Flag{\dynkin C{xo}}{\dynkin A{x}}
%\Flag{\dynkin C{xoo}}{\dynkin A{x}}
%\Flag{\dynkin C{xo.ooo}}{\dynkin{A}{x}}%
%\Flag{\dynkin C{*xo}}{\dynkin C{*x}}
%\Flag{\dynkin C{**xo}}{\dynkin C{**x}}
%\Flag{\dynkin C{*.*xo}}{\begin{dynkinDiagram}{C}{**.*x}\dynkinBrace[r-1]{1}{4}\end{dynkinDiagram}}
\LastFlag{1}{C_r}{%

@@

@@

% D's

@@

\Flag{\dynkin{D}{oxo.oooo}}{\dynkin{A}{x}}

@@

For a flag variety \(X=G/P\), the root system is also \emph{\(P\)-graded} by \(\sum n_i\), but summing only over the noncompact (crossed) simple roots \cite{Cap/Slovak:2009} p.~292.
The \(P\)-grade is also called the \emph{\(P\)-height} \cite{Cap/Slovak:2009} p.~292.
A root is \emph{\(P\)-maximal}, also called a \emph{box root}, if it has maximal \(P\)-grade in its irreducible factor.
The \emph{box} is the set of box roots, terminology which roughly follows what \cite{Buch.Chaput.Mihalcea.Perrin:2018}, \cite{Lam.Williams:2008} p.~57 might perhaps call the \emph{maximal box}, by analogy with Young tableaux.
We will see in lemma~\vref{lemma:box.roots} that the box generates the root system of \(\breve{G}\).

@@

The unipotent radical \(G_+\subseteq P\) has Lie algebra \(\LieG_+\subseteq\LieP\) the sum of the root vectors of the positively graded roots \cite{Borel:1991} p.~197 Proposition~14.18, \cite{Knapp:2002} p.~482.

@@

The sum of the Cartan subalgebra with the sum of the root spaces of the zero graded roots is the semisimple factor of the reductive Levi factor \(\LieGZ\) of \(\LieP\) \cite{Borel:1991} p.~197 Proposition~14.18.

@@

We can and will pick the hyperplane to be the zero locus of a real linear function taking on value zero on uncrossed simple roots and one on crossed simple roots \cite{Cap/Slovak:2009} p.~239 Proposition~3.1.2.

@@

Under the heading  \(\breve\LieG\), we draw the roots of the symmetry Lie algebra of the associated cominuscule, and under the heading \(\LieG'\) the roots of the automorphism Lie algebra.

@@

Similarly, the associated cominuscule of any adjoint variety is \(\dynkin[parabolic=1]{A}{1}=(\Proj{1},\PSL{2})\), since the adjoint varieties are precisely those with invariant contact structures, which arise from a hyperplane in each tangent space, invariant under the parabolic, hence a sum of root spaces, so a single root, negative for the parabolic, doesn't have its root vector in this hyperplane, i.e. a single root is at higher weight.
So the associated cominuscule is complementary to the hyperplane, one dimensional, hence a projective line.

@@

Recall that every root system has a unique highest root, highest in the sense of the positive roots, i.e. the highest weight of the adjoint representation \cite{Serre:2001} p.~61 Theorem~3.
Every root can be brought to the highest root by successively adding simple roots, passing through a sequence of roots \cite{Fulton/Harris:1991} pp.~330--331 \S{21.3}, \cite{Humphreys:1978} p.~56, \cite{Knutson:2022}, \cite{Serre:2001} p.~46 2.12, \cite{Serre:2001} p.~58 Theorem~3.
As this process can only raise the \(P\)-height, the highest root is \(P\)-maximal.
Every \(P\)-maximal root can thus be brought to the highest root by adding only \(P\)-compact simple roots.
So the box becomes a connected graph, with the highest root as one vertex.
The vector space \(\LieZ'\) is therefore an irreducible \(\GZ\)-module \cite{Knapp:2002} p.~332 proposition~5.105.

Pick a root vector \(e_{\aa}\in\LieZ\) for a root \(\aa\in B\).
We need to prove that \(\aa\in B'\) i.e. \(\aa\) is a box root, i.e. \(P\)-maximal.
Suppose that \(\aa+\bb\) is a root, for some root \(\bb\) of positive \(P\)-height.
So then \(0\ne\lb{e_{\aa}}{e_{\bb}}\in\gr[\aa+\bb]\LieG\) for some root vector \(e_{\bb}\) of \(\bb\) \cite{Serre:2001} p.~46 2.12.
But then \(e_{\aa}\) is not in \(\LieZ\), i.e. does not centralize \(\LieG_+\).
But the highest root is \(\aa\) or is of the form \(\aa+\bb\) for some positive root \(\bb\).
Hence \(\bb\) is \(P\)-compact, i.e. \(\aa\) is \(P\)-maximal.
So \(\aa\) is a box root: \(\aa\in B'\).

@@

By Langlands decomposition \cite{Knapp:2002} p.~482,  \(G,P,X,G_+,G_0\) are connected.

@@

The exponential map on the Lie algebra of a unipotent linear algebra group is a biholomorphism mapping subalgebras to subgroups, so taking \(\LieZ\to Z\), an isomorphism of schemes \cite{Milne:2017} p.~289 Proposition 14.32.
%In a Chevalley basis, each element
%g=e^{\sum t_{\alpha} e_{\alpha}}\in G_+
%acts on any element \(e_{\beta}\in G_+\) by
%\Ad_g e_{\beta}.
%This expands out into a sum of terms in \(t_{\alpha}\) with entirely positive coefficients unless \(\log g\) is a sum of \(P\)-maximal roots.
%\todo{Justify that last sentence.
%It might be related to Demazure \& Gabriel, p.~499.}
%So the center \(Z=Z_{G_+}\) of \(G_+\) consists precisely of the exponentials of elements of
%\LieZ=\bigoplus\LieG_{\alpha},
%the sum being over the box, i.e. over the \(P\)-maximal roots, so \(Z\) is connected.
By definition of \(\breve{G}\), \(\breve{G}:=\left<Z,aZ\right>\) is a connected complex Lie group, since \(Z\) is connected. 
Moreover, \(\breve{G}\) is connected in the algebraic Zariski topology by Chevalley's theorem on constructible sets \cite{Conrad:2017} p.~38 Proposition 16.2.1.

@@

%Two associated cominuscule subvarieties are tangent at a point just when they are equal.

@@

The vector spaces \(\LieZ\) and \(a\LieZ\) are irreducible \(\GZ\)-modules \cite{Knapp:2002} p.~332 proposition 5.105.
Hence \(\breve\LieG\) is a \(\GZ\)-module.
As \(\GZ\) is reductive, \(\breve\LieG\) is a direct sum of irreducible \(\GZ\)-modules:
\[
\breve\LieG=\LieZ\oplus{\breve\LieG}_0\oplus a\LieZ.
\]

Suppose that \(\aa\) is a \(P\)-compact root, i.e. a root of \(\GZ\).
Reflection in \(\aa\) is carried out by an element of the Weyl group of \(\GZ\), and so preserves the \(P\)-grading.
So if \(\bb\) is any \(P\)-maximal root, then so is its \(\aa\)-reflection.
In other words, reflection in \(\aa\) preserves the box.
Recall that an \emph{\(\aa\)-root string} is a directed path of roots, each given by adding \(\aa\) to the previous.
Reflection in \(\aa\) moves \(\bb\) along an \(\aa\)-root string.

@@

By definition \(P'\) preserves \(\breve\LieG+\LieP\).
Pick a root \(\mx\bb\) in the box, not perpendicular to our \(P\)-submaximal \(P\)-positive root \(\aa\).
Note that if two roots have more than a right angle between them, then their sum is a root and their root vectors have nonzero Lie bracket \cite{Serre:2001} p.~29.
Suppose that \(\aa\) is \(P\)-positive.
The root system generated by \(\aa\) and \(\mn\bb:=-\mx\bb\) is of rank \(2\), and our \(P\)-grading grades that rank \(2\) root system, so we can examine in our pictures above every possible case.
We see that the root vectors of \(\aa\) and \(\mn\bb\) have more than a right angle between them and that their sum is a root.
Since their \(P\)-grading is at least that of \(\mn\bb\), \(P\)-negative, the Lie bracket of their root vectors is not in \(\LieP\), but also is not \(P\)-minimal, so not in \(\breve\LieG\).
Hence their root vectors bracket out of \(\breve\LieG+\LieP\), 
Hence the Lie algebra \(\LieG'\) of \(G'\) contains no such.

@@

The \emph{Hasse diagram} draws dots on the plane, one for each positive root and a line from each root upward to each of its successors, labelled by the number of the simple root by which they differ.

@@

The Hasse diagram of \(C_4\) is

@@

Above, we see that the labelling of \(E_8\) is
\dynkin[labels*={1,,,,,,,}]E8
Take the \(E_8\)-flag variety with Dynkin diagram
